---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 120)
```

# bootabc

<!-- badges: start -->
[![R-CMD-check](https://github.com/nivr/bootabc/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nivr/bootabc/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

bootabc computes bootstrap confidence intervals for the KPIs of an A/B test, from data
aggregated to one row per customer. It handles compound ratio KPIs (such as spend per
active day), variant comparisons (lift and difference), CUPED variance reduction, and BCa
intervals -- and it runs the resampling through a streaming C++ kernel, so a
million-customer experiment bootstraps in seconds.

Every supported KPI is a ratio of per-customer column sums, so the bootstrap resamples
those **sums** rather than re-evaluating the statistic on each resample. That makes the
engine exact (no normal or sqrt(N) approximation) and fast, and lets KPIs be written in a
small, safe grammar instead of parsed from free-form strings.

## Installation

``` r
# requires a C++ toolchain (Rtools on Windows)
# install.packages("remotes")
remotes::install_github("nivr/bootabc")
```

## Quick start

```{r quickstart, message = FALSE}
library(bootabc)
library(dplyr)

set.seed(1)
customers <- group_by(
  data.frame(
    spend       = rgamma(4000, shape = 2),
    active_days = rpois(4000, 4) + 1,
    variant     = sample(c("control", "treatment"), 4000, replace = TRUE)
  ),
  variant
)

result <- bootstrap_measures(
  customers,
  arpu          = mean(spend),                   # a simple mean
  spend_per_day = sum(spend) / sum(active_days), # a compound ratio KPI
  comparison = variant, reference = "control",
  iterations = 2000, seed = 1
)

result

confidence_intervals(result, probs = c(lower = 0.025, upper = 0.975))
```

A `ratio` interval that straddles 1 (or a `difference` that straddles 0) means the effect
is not distinguishable from noise -- as here, where the data is null.

## The measure grammar

KPIs are built from `sum()`, `mean()`, `weighted.mean()`, and division:

``` r
bootstrap_measures(
  customers,
  revenue       = sum(spend),
  arpu          = mean(spend),
  spend_per_day = sum(spend) / sum(active_days),
  weighted      = weighted.mean(spend, active_days)
)
```

Anything outside the grammar -- `median()`, `quantile()`, a bare column, arithmetic other
than `/` -- is rejected, because only ratios of sums have an exact sum-resampling
bootstrap.

## Comparisons and strata

Name a `comparison` column to get within-stratum lift (ratio) and difference
distributions; pass a `reference` level to compare each arm against it (otherwise all
pairs are formed). Group by extra columns to estimate each stratum separately:

``` r
bootstrap_measures(
  group_by(customers, country, variant),
  arpu = mean(spend),
  comparison = variant, reference = "control"
)
```

## CUPED variance reduction

With a pre-experiment covariate, de-noise the metric before bootstrapping -- the point
estimate is unchanged in expectation while the interval narrows:

``` r
adjusted <- cuped(customers, spend = spend_pre)
bootstrap_measures(adjusted, arpu = mean(spend), comparison = variant, reference = "control")
```

## Interval methods

`confidence_intervals(result, method = ...)` supports:

- `"percentile"` (default) -- quantiles of the draws
- `"basic"` -- those quantiles reflected about the observed estimate
- `"bca"` -- bias-corrected and accelerated, from the per-customer jackknife (a two-sample
  jackknife for comparisons)

## Plotting

```{r forest, fig.width = 7, fig.height = 3, dpi = 120, fig.cap = "Lift (ratio) and difference for each KPI, with 95% intervals."}
plot_intervals(confidence_intervals(result))
```

## Performance

Exact resampling is unavoidably `O(iterations x customers x columns)`, but the streaming
kernel keeps memory at `O(columns)` per iteration and runs in C++. Single-threaded timings
(indicative; hardware varies):

| customers | iterations | KPIs |  time |
|----------:|-----------:|-----:|------:|
|   100,000 |     10,000 |    2 |  2.3s |
| 1,000,000 |     10,000 |    2 | 26.4s |

``` r
n <- 1e6
data <- group_by(data.frame(
  spend = rgamma(n, 2), active_days = rpois(n, 4) + 1,
  variant = sample(c("control", "treatment"), n, replace = TRUE)
), variant)

system.time(bootstrap_measures(
  data, arpu = mean(spend), spend_per_day = sum(spend) / sum(active_days),
  comparison = variant, reference = "control", iterations = 10000, seed = 1
))
```

## License

MIT.
